\documentstyle[12pt]{article}

\oddsidemargin -.5in
\textwidth 7.5in
\textheight 9in
\topmargin -30pt

\begin{document}

%\setcounter{page}{0}


\begin{center}
  67.612  Combinatorial Optimization and Integer Programming, Spring 93.
\end{center}

\begin{center}
  Midterm Exam, Monday, March 29, 1993.
\end{center}

Please do all four problems. Show all work. No books or calculators allowed.
You may use any result from class, the homeworks, or the texts.
You may use one sheet of handwritten notes.
The exam lasts two hours.


\begin{enumerate}
   \item (20 points)
      Find the shortest path from vertex 1 to vertex 6 in the following
      graph. Edge lengths are indicated on the edges.
   \item (20 points)
      For each of the following graphs, find the maximum cardinality matching
      and give a short proof that it realy does contain the largest possible
      number of edges.
   \item (40 points)
      \begin{enumerate}
         \item (35 points; each part is worth five points.) \label{particular}
            Consider the binary knapsack problem:
            \begin{displaymath}
              \begin{array}{lrcrcrcrcrcl}
     \max        & 10x_1 & + & 8x_2 & + & 5x_3 & + & 2x_4 & + & x_5 \\
     \mbox{st }  &  2x_1 & + & 4x_2 & + & 5x_3 & + & 3x_4 & + & 4x_5 & \leq & 10
                                 \qquad \qquad (KP)  \\
                 &&&&&&&&& x_i && \mbox{binary.}
              \end{array}
            \end{displaymath}
            Let $P=\{x \in {\Re}^5_+:2x_1+4x_2+5x_3+3x_4+4x_5\leq 10,
                    0 \leq x_i \leq 1\}$ and~$S=P\cap{\cal Z}^5$.
            \begin{enumerate}
              \item
                 Solve the LP-relaxation
                 $\max\{10x_1+8x_2+5x_3+2x_4+x_5 : x \in P \}$.
              \item
                 What is the dimension of $S$?
              \item
                 Prove that $x_1+x_2+x_3 \leq 2$ is a valid inequality for~$S$.
              \item
                 Prove that $x_1+x_2+x_3 \leq 2$ is a facet of conv($S$),
                 the convex hull of~$S$.
              \item
                 What is the Chvatal-Gomory rank of the inequality
                 $x_1+x_2+x_3 \leq 2$?
              \item
                 Prove that $x_3+x_4+x_5 \leq 2$ is a valid inequality for~$S$.
              \item
                 Show that the inequality $x_3+x_4+x_5 \leq 2$ can be lifted
                 to give a stronger valid inequality for~$S$.
            \end{enumerate}
         \item (5 points)
            Consider the general binary knapsack problem:
            \begin{displaymath}
              \begin{array}{lrcl}
                \max          & \sum_{i=1}^n c_ix_i \\
                \mbox{s.t. }  & \sum_{i=1}^n a_ix_i & \leq & b \\
                              &            x_i  && \mbox{binary.}
              \end{array}
            \end{displaymath}
            Generalize the results of part~\ref{particular} to give a class of
            valid inequalities for this problem.
      \end{enumerate}
   \item (20 points) The {\em max clique} feasibility problem can be described
     as follows:
     \begin{quote}
       An {\em instance} consists of a graph $G=(V,E)$ and an integer $k$.
       The answer is {\em YES} if the graph contains a clique with at
       least $k$ vertices; otherwise, the answer is~{\em NO}.
     \end{quote}
     Show that the {\em max clique} problem is $NP$-complete.
     (Hint: You may want to look at the {\em complement} of $G=(V,E)$,
     that is, the graph $G=(V,E')$ where $e$ is in $E'$ if and only if
     it is not in~$E$.)
\end{enumerate}

\end{document}
