\documentstyle[12pt]{article}

\oddsidemargin -.5in
\textwidth 7.5in
\textheight 9in
\topmargin -30pt

\begin{document}

\setcounter{page}{0}

\begin{tabular}{cl}
  \hspace{5in} & Name:
\end{tabular}

\begin{center}
  67.612  Combinatorial Optimization and Integer Programming, Spring 91.
\end{center}

\begin{center}
  Midterm Exam, Tuesday, March 5, 1991.
\end{center}

Please do all five problems. Show all work. No books or calculators allowed.
You may use any result from class, the homeworks, or the texts.

\vspace{2in}

\begin{center}
\begin{tabular}{c|@{\hspace*{1in}}c}
  Q1 & \qquad \\ \hline
  Q2 & \qquad \\ \hline
  Q3 & \qquad \\ \hline
  Q4 & \qquad \\ \hline
  Q5 & \qquad \\ \hline \hline
  Total &
\end{tabular}
\end{center}

\pagebreak

\begin{enumerate}
   \item (20 points)
     Let $G=(V,E)$ be a bipartite graph with $(W_1,W_2)$ being a
     bipartition of $V$. Assume $\mid W_1\mid =\mid W_2\mid =n$.
     Show that $G$ has a matching of cardinality $n$ if
     \begin{displaymath}
       \mid \{ \mbox{neighbours of } S \} \mid =
       \mid \{u\in W_2: (v,u) \in E \mbox{ for some } v \in S \}\mid  \; \geq \;
               \mid S\mid
     \end{displaymath}
     for all $S \subseteq W_1$.
   \pagebreak
   \item (15 points)  Show that the graph feasibility problem
     \begin{quote}
       Is $G$ bipartite?
     \end{quote}
     is in both $NP$ and $CoNP$.
   \pagebreak
   \item (25 points)  Consider the graph
     \vspace{1in}
     \begin{enumerate}
       \item (5 points) Give a maximum cardinality matching for this graph.
       \item (5 points) \label{formIP}
         Formulate the problem of finding the maximum cardinality
         matching in the form
         $\max\{c^Tx : Ax \leq b, x \geq 0, x \mbox{ integer} \}$.
       \item (5 points) Give an optimal solution
         to the linear programming relaxation of the integer
         program given in part~\ref{formIP}.
       \item (5 points) Suggest an additional constraint for the LP relaxation.
       \item (5 points) Formulate and solve the dual of the LP relaxation with
         the additional constraint.
         (The dual of the problem $\max\{c^Tx : Ax \leq b, x \geq 0 \}$ is
         $\min\{b^Ty : A^Ty \geq c, y \geq 0 \}$.)
     \end{enumerate}
   \pagebreak
   \item (15 points)  Let $G=(V,E)$ be a graph with weights $w_e$ on the edges.
     Assume $w_{e_i}\neq w_{e_j}$ for any pair of edges $e_i$ and $e_j$.
     Let $T$ be a minimum weight
     spanning tree on this graph.  Divide the vertices into two sets
     $U$ and $V\setminus U$. Show that $T$ must contain the shortest
     edge between $U$ and $V \setminus U$.
   \pagebreak
   \item (25 points. Each part is worth five points.)
     Give short justifications for your answers.
     \begin{enumerate}
       \item An instance $d$ of a feasibility problem $X \in NP$
         depends upon two positive integer parameters $m$ and $n$.
         Assume $d$ requires storage $m2^n$ in binary, and that we
         know an algorithm $A$ which solves $d$ in time $2^{m+n}$.
         \begin{enumerate}
           \item Can we conclude $X$ is in $P$?
             \vspace*{1in}
           \item Can we conclude $X$ is not in $P$?
             \vspace*{1in}
           \item Assume we know in addition that $m\leq n$ for every
             instance $d$ of $X$.  What can we conclude now?
             \vspace*{1in}
         \end{enumerate}
       \item
         \begin{enumerate}
           \item Is the given $(s,t)$ flow optimal? (The numbers on the
             arcs indicate the capacity and the flow.)
             \vspace*{1.8in}
           \item Is the given matching optimal?
         \end{enumerate}
     \end{enumerate}
\end{enumerate}

\end{document}
