\documentclass[12pt]{article}

\begin{document}

% My definitions of page layout
%\oddsidemargin .3in
%\textwidth 6in
%\textheight 8.8in
%\topmargin -30pt

\pagestyle{myheadings}
\markboth{Stochastic programming papers}{Stochastic programming papers}

\begin{center}
  \begin{large}
    {\bf MATP6960 Stochastic Programming} \\
    {\bf Presentations} \\   {\bf Fall 2002} 
  \end{large}
\end{center}

\section{Schedule}

Student presentations will occur on {\bf December 11}, from 6.30pm to 9.30pm.
Each talk will last approximately fifteen to twenty minutes.
The order of presentation will be random.

You should describe the major contribution(s) of the paper you present.
You should type up a short (3-5 pages) analysis of the paper(s),
which I will photocopy and give out
{\em before} your presentation. Thus, please give me your summary by December 10.

In order to encourage questions, your grade will not be lowered if you are unable to answer questions from other students, but it may be raised. Moreover, I may give some
{\em bonus points} for asking a particularly good question.


\section{Papers on reserve}


The online version of this document will contain pointers to many of the papers.

\section{Papers for your presentations}

This is a list of papers that either discuss some of the topics I covered
in more detail or discuss extensions of the material.
You can pick a paper from this list to present, or you can suggest
a different paper, or I will suggest a paper.
In some cases, it will be appropriate to present two papers.
I have copies of all of these papers, and you can find many of them
either in the library or online.

\subsection{Surveys}

\begin{itemize}
\item
Sen and Higle~\cite{sen1}
\item
Wets~\cite{wets1}
\end{itemize}

 \subsection{Computation}

\begin{itemize}
\item
Birge~\cite{birge4}
\item
Fragniere, Gondzio, and Vial~\cite{fragniere2}
\item
Linderoth and Wright~\cite{linderoth3}
\end{itemize}

\subsection{Stochastic integer programs}

\begin{itemize}
%\item
%Ahmed, King, and Parija~\cite{ahmed1}
\item
Car\oe  \ and Schultz~\cite{caroe2}
\item
Car\oe  \ and Tind~\cite{caroe1}
\item
Takriti and Birge~\cite{takriti1}
\end{itemize}

\subsection{Sampling methods}

\begin{itemize}
\item
Godfrey and Powell~\cite{godfrey1}
\item
Linderoth, Shapiro, and Wright~\cite{linderoth4}
\item
Rush, Mulvey, Mitchell, and Willemain~\cite{mulvey3}
\item
Verweij, Ahmed, Kleywegt, Shapiro, and Nemhauser~\cite{verweij1}
\end{itemize}

\subsection{Robust optimization}

\begin{itemize}
\item
Mulvey, Vanderbei, and Zenios~\cite{mulvey1} ({\bf Lee})
\item
Takriti and Ahmed~\cite{takriti2}
\end{itemize}

\subsection{Multistage problems}

\begin{itemize}
%\item
%Berger et al~\cite{berger1}
\item
Birge et al~\cite{birge2}
\end{itemize}

\subsection{Bounding techniques}

\begin{itemize}
\item
Mak, Morton, and Wood~\cite{mak1}
\end{itemize}

\subsection{Variance reduction}

\begin{itemize}
\item
Higle~\cite{higle4}
\end{itemize}

\subsection{Financial applications}

\begin{itemize}
\item
Drijver, Klein Haneveld, and van der Vleck~\cite{drijver1}
\item
Fleten, H{\o}yland and Wallace~\cite{fleten1}
\item
Gondzio and Kouwenberg~\cite{gondzio8}
\item
Gulpinar, Rustem, and Settergren~\cite{gulpinar1}
\item
Haugen and Wallace~\cite{haugen1}
\item
Mulvey and Simsek~\cite{mulvey6}
\item
Seshadri et al~\cite{seshadri1}
\item
Zhao and Ziemba~\cite{yzhao1}
\end{itemize}

\subsection{Applications}

\begin{itemize}
\item
Cormican, Morton, and Wood~\cite{cormican1}
\end{itemize}

\subsection{Test problems}

\begin{itemize}
\item
Ariyawansa and Felt~\cite{ariyawansa2}
\end{itemize}

\bibliography{optim}
\bibliographystyle{plain}

\end{document}
