\documentclass[12pt]{article}

\pagestyle{empty}

\oddsidemargin -.5in
\textwidth 7.5in
\textheight 10in
\topmargin -30pt
\headsep 0in
\headheight 0in

\newcommand{\re}{I \!\! R}

\begin{document}

\begin{center}
  \begin{large}
     MATP6640/DSES6770 Linear Programming, Homework 4.
  \end{large}
\end{center}

\begin{flushright}
   Due:  Tuesday, March 28, 2006.
\end{flushright}


%\vspace{\baselineskip}

\begin{enumerate}
\item
Use the network simplex algorithm to find the minimum cost flow for the problem
with
the following linear programming representation:
\begin{displaymath}
\begin{array}{lrcrcrcrcrcrcl}
\min & & - & 2x_{13}& - & x_{23} & - & 2x_{24} &&& - & 10 x_{41} \\
\mbox{subject to} & -x_{12} & - & x_{13} &&&&&&& + & x_{41} & = & 0 \\
& x_{12} &&& - & x_{23} & - & x_{24} &&&&& = & 0 \\
&&& x_{13} & + & x_{23} &&& - & x_{34} &&& = & 0 \\
&&&&&&& x_{24} & + & x_{34} & - & x_{41} & = & 0 \\
& \multicolumn{11}{r}{0 \, \leq \, x_{12}, x_{13}, x_{23}, x_{24}, x_{34}}
& \leq & 1 \\
&&&&&&&&& 0 & \leq & x_{41} & \leq & 10
\end{array}
\end{displaymath}
Use the initial basic feasible solution with basic variables
$ %\begin{displaymath}
x_{12} = 1, x_{34}=1, x_{41}=1
$ %\end{displaymath}
and nonbasic variables
$ %\begin{displaymath}
x_{13}=0, x_{23}=1, x_{24}=0.
$ %\end{displaymath}
You should need three iterations.
\item   \label{q.multiform}
A multicommodity network flow problem on a graph $G=(V,E)$ with $K$ commodities
can be expressed in a path-based formulation, with a set of paths $P(k)$
for each commodity~$k$.
Assume each commodity has a single source node and a single sink node.
This leads to the linear program
\begin{displaymath}
\begin{array}{lrcll}
\min & \sum_{k=1}^K \sum_{p \in P(k)} w_p f_p \\
\mbox{subject to} & \sum_{p \in P(k)} f_p & = & d_k & k = 1,\ldots,K  \\
& \sum_{k=1}^K \sum_{p \in P(k): e \in p} f_p & \leq & u_e & \forall e \in E  \\
& f_p & \geq & 0 & \forall p \in P(k), \mbox{ for } k=1,\ldots,K
\end{array}
\end{displaymath}
where $d_k$ is the demand for commodity $k$,
$u_e$ is the capacity of edge~$e$,
and $w_p$ is the cost of sending one unit of flow along path~$p$.
Let $c_e^k$ be the cost of shipping one unit of commodity $k$ on arc~$e$;
we have $w_p=\sum_{e \in p} c_e^k$ for $p \in P(k)$.
The set $P(k)$ typically does not include all paths that can be used by commodity~$k$.
Assume the LP is solved, and you know a primal and dual optimal solution.
Show that determining whether another path should be added to one of the
sets $P(k)$ requires solving a shortest path problem.

\item
Consider a multicommodity network flow problem with three commodities
on a graph $G=(V,E)$ with five nodes and seven edges.
The seven edges and their costs and capacities are as follows:
\begin{displaymath}
\begin{array}{l|rrrrrrr}
\mbox{edge} & (1,2) & (1,5) & (2,3) & (2,4) & (2,5) & (3,4) & (4,5) \\ \hline
\mbox{cost}   &  10  &  11 &  7 & 10 & 6 & 6 & 8 \\
\mbox{capacity} & 30 & 40 & 25 & 50 & 40 & 25 & 50
\end{array}
\end{displaymath}
The costs are identical for each commodity.
25 units of commodity A need to be moved from node 1 to node~3.
20 units of commodity B need to be moved from node 5 to node~3.
20 units of commodity C need to be moved from node 4 to node~1.
The current set of paths consists of $1 \rightarrow 2 \rightarrow 3$
and $1 \rightarrow 5 \rightarrow 4 \rightarrow 3$ for commodity~A,
$5 \rightarrow 2 \rightarrow 3$ for commodity~B,
and $4 \rightarrow 5 \rightarrow 1$ for commodity~$C$.
Use the methodology of Question~\ref{q.multiform} to show
that this is not an optimal set of paths.
What is the optimal solution?
\item
Chvatal, Question 15.8 on page 238.
\end{enumerate}


\vfill

\begin{tabular}{@{\hspace{.5in}}l}
   John Mitchell  \\
   Amos Eaton 325  \\
   x6915.  \\
   mitchj@rpi.edu  \\
   Office hours: Tuesday: 2 -- 3.30pm.
\end{tabular}

\end{document}
